title: "nyhousing_wrangling"
author: "Robbie Zielinski"
date: "2/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RMySQL)
library(ggmap) 
library(operators)
library(tidyverse)
library(pbapply)
library(tcltk)
library(lubridate)
library(doParallel)
```


```{r}
db <- dbConnect(MySQL(), user = 'stat231', password = 'stat231project', dbname = "nyhousing", host = '148.85.253.214')
transactions <- tbl(db, "transactions")
transactions_df <- dbReadTable(db, "transactions")
```

```{r}
transactions <- transactions %>%
  mutate(address_full = CONCAT(Address, ", ", Town, ", ", "NY"))

transactions_df <- transactions_df %>%
  mutate(address_full = paste(Address, Town, "NY", sep = ", "))

#small <- transactions %>%
 # filter(Town == 'Albany')
#transactions2 <- transactions %>%
#  filter(Town == "Rochester")
```

```{r}
no_cores <- detectCores() - 1
registerDoParallel(cores=no_cores)  
cluster <- makeCluster(no_cores, type="FORK") 
#full_address <- transactions %>%
#  select(address_full) %>%
#  collect() %>%
#  unlist() %>%
#  as.vector()

full_address <- transactions_df %>%
  select(address_full) %>%
  unlist() %>%
  as.vector()

geocodes <- pblapply(full_address, function(x) tryCatch(geocode(x, output = "latlon", messaging = FALSE, source = "dsk"), error = function(e) data.frame(lon = NA, lat = NA)), cl = cluster)  
stopCluster(cluster)
```

```{r}
full_address <- data.frame(full_address)
names(full_address) <- c("address_full")
geocodes <- unlist(geocodes) %>%
  matrix(ncol = 2, byrow = TRUE) %>%
  data.frame
names(geocodes) <- c("lon", "lat")
geocodes <- bind_cols(full_address, geocodes) %>%
  distinct()
geocodes$address_full <- as.character(geocodes$address_full)
transactions2 <- transactions_df %>%
  left_join(geocodes, by = "address_full")
```

```
pb <- tkProgressBar(title = "progress bar", min = 0,
                    max = nrow(transactions), width = 300)

for(i in 1:nrow(transactions2)) { ### Allows me to continue with errors
  result <- tryCatch(geocode(transactions2$address_full[i], output = "latlon", source = "dsk"),
                     error = function(e) data.frame(lon = NA, lat = NA, address = NA))
  transactions2$lon[i] <- as.numeric(result[1])
  transactions2$lat[i] <- as.numeric(result[2])
  setTkProgressBar(pb, i, label=paste( round(i/nrow(transactions2)*100, 0),
                                        "% done"))
}

close(pb)
```


``{r}
transactions2 <- transactions2 %>%
  mutate(SalePrice = REPLACE(SalePrice, ',', ''),
         month = SUBSTRING_INDEX(SaleDate, '/', 1),
         day = SUBSTRING_INDEX(SUBSTRING_INDEX(SaleDate, '/', 2), '/', -1),
         year = SUBSTRING_INDEX(SaleDate, '/', -1))
```
```{r}
counties <- c("Albany", "Allegany", "Broome", "Cattaraugus", "Cayuga", 
              "Chautauqua", "Chemung", "Chenango", "Clinton", "Columbia",
              "Cortland", "Delaware", "Dutchess", "Erie", "Essex", "Franklin",
              "Fulton", "Genesee", "Greene", "Hamilton", "Herkimer", "Jefferson",
              "Lewis", "Livingston", "Madison", "Monroe", "Montgomery",
              "Nassau", "Niagara", "Oneida", "Onondaga", "Ontario", "Orange",
              "Orleans", "Oswego", "Otsego", "Putnam", "Rensselaer", "Rockland",
              "St. Lawrence", "Saratoga", "Schenectady", "Schoharie",
              "Schuyler", "Seneca", "Steuben", "Suffolk", "Sullivan",
              "Tioga", "Tompkins", "Ulster", "Warren", "Washington", "Wayne",
              "Westchester", "Wyoming", "Yates")
              
long_min <- c(-74.3, -78.3, -76.2, -79.1, -76.8, -79.8, -77.0, -75.9, -74.0, -74.0,
              -76.3, -75.5, -74.0, -79.2, -74.4, -74.8, -74.8, -78.5, -74.6, -74.9,
              -75.3, -76.5, -75.9, -78.1, -76.0, -78.0, -74.8, -73.8, -79.1, -75.9,
              -76.5, -77.7, -74.8, -78.5, -76.7, -75.5, -74.0, -73.8, -74.3, -75.9, 
              -74.2, -74.4, -74.8, -77.1, -77.0, -77.8, -73.5, -75.2, -76.6, -74.8,
              -74.3, -73.7, -77.4, -74.0, -78.5, -77.4)

long_max <- c(-73.6, -77.7, -75.3, -78.3, -76.2, -79.0, -76.5, -75.3, -73.3, -73.3,
              -75.8, -74.4, -73.4, -78.4, -73.3, -73.9, -74.1, -77.9, -73.7, -74.0,
              -74.6, -75.4, -75.1, -77.4, -75.2, -77.3, -74.0, -73.4, -78.4, -75.0,
              -75.8, -76.9, -73.9, -77.9, -75.7, -74.6, -73.5, -73.2, -73.9, -74.5,
              -73.5, -73.8, -74.1, -76.6, -76.6, -76.9, -71.8, -74.3, -76.0, -73.9, 
              -73.4, -73.2, -76.7, -73.4, -77.9, -76.9)

lat_min <- c(42.4, 42.0, 42.0, 42.0, 42.6, 42.0, 42.0, 42.2, 44.4, 41.9, 42.4, 41.8,
             41.4, 42.4, 43.7, 44.0, 42.9, 42.8, 42.0, 43.2, 42.8, 43.6, 43.4, 42.4,
             42.7, 42.9, 42.7, 40.5, 43.0, 42.8, 42.7, 42.5, 41.1, 43.1, 43.1, 42.3,
             41.3, 42.4, 40.9, 44.0, 42.7, 42.7, 42.3, 42.2, 42.5, 42.0, 40.6, 41.4,
             42.0, 41.5, 43.2, 42.9, 43.0, 40.8, 42.5, 42.4)

lat_max <- c(42.9, 42.6, 42.5, 42.6, 43.5, 42.6, 42.3, 42.8, 45.0, 42.5, 42.8, 42.6,
             42.1, 43.1, 44.6, 45.0, 43.3, 43.2, 42.5, 44.2, 44.1, 44.4, 44.3, 43.0,
             43.2, 43.4, 43.1, 41.0, 43.4, 43.6, 43.3, 43.0, 41.7, 43.4, 43.7, 42.9,
             41.6, 43.0, 41.4, 45.1, 43.4, 43.0, 42.9, 42.6, 43.1, 42.6, 41.3, 42.0,
             42.5, 42.2, 43.8, 43.8, 43.4, 41.4, 42.9, 42.8) 
             
counties_lat_lon <- data.frame(counties, long_min, long_max, lat_min, lat_max)
```


```{r}
transactions2 <- transactions2 %>%
  mutate(SalePrice = gsub(",", "", SalePrice),
         FullSaleDate = SaleDate) %>%
  separate(FullSaleDate, c("month", "day", "year"), sep = "/")
```

```{r}
#transactions2 <- collect(transactions2)
dbWriteTable(db, "transactions2", transactions2, overwrite = TRUE)
```